Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-10-25 algorithm.tex
AA

\comment{TODO AND: c'e' un problema col mathcal, dai un'occhiata qui sotto}
Each iteration is called \emph{era}. The $i$-th era is characterized by: 
	\begin{itemize}
	\item the set of regions $\mathcal{R}_{i}$ in which the configuration space
	$mathcal{C}(S)=V_{1}\times\dots\times V_{n}$ is partitioned
	\item an approximation $\mathscr{P}_{i}$ of the Pareto-optimal set.
	\item a set $test_{i}$ of at most $K$ configurations evaluated
	in the era.
	\end{itemize}
$K$ is a parameter that must be set before the algorithm runs.
Note that $\mathcal{R}_{i}$ is a partition of the configuration space, i.e.
	\begin{align}\begin{array}{l}
		\bigcup_{R\in\mathcal{R}_i} R = \mathcal{C}(S) \mbox{ and}\\
		R' \cap R'' = \emptyset, \forall R',R'' \in \mathcal{R}_i
	\end{array}\end{align}

We now describe formally PS.

>>>>

Each iteration is called \emph{era}. The $i$-th era is characterized by: 
	\begin{itemize}
	\item the set of regions $\mathcal{R}_{i}$ in which the configuration space
	$\mathcal{C}(S)=V_{1}\times\dots\times V_{n}$ is partitioned
	\item an approximation $\mathscr{P}_{i}$ of the Pareto-optimal set.
	\item a set $test_{i}$ of at most $K$ configurations evaluated
	in the era.
	\end{itemize}
$K$ is a parameter that must be set before the algorithm runs.
Note that $\mathcal{R}_{i}$ is a partition of the configuration space, i.e.
	\begin{align}\begin{array}{l}
		\bigcup_{R\in\mathcal{R}_i} R = \mathcal{C}(S) \mbox{ and}\\
		R' \cap R'' = \emptyset, \forall R',R'' \in \mathcal{R}_i
	\end{array}\end{align}

We now formally describe PS.

### 2014-10-25 experiments.tex
AA

\section{A Case Study: VLIW Hardware/Software Design Space}
\secL{ee}
In order to evaluate the behavior of the proposed approach
we use a parameterized Very Long Instruction
Word (VLIW) architecture~\cite{kathail_tr00} as testbed. The philosophy
behind a VLIW system makes this a natural choice for several reasons:
\begin{itemize}
\item \emph{Multi-Objective trade-offs}: VLIW architectures allow designers
to trade between power, energy and performances objectives, resulting
in extended Pareto sets which are an ideal ground for our testing
purposes focused on parameter-space representation
\item \emph{Compiler awareness}: code scheduling for the execution of
applications is statically obtained by the compiler which
performs code transformations ``being aware'' of the underlying hardware
configuration.  This tight hardware/software dependence makes the
VLIW scenario perfectly suitable for investigating hardly predictable
parameter interactions.
\end{itemize}

@@

\subsection{Evaluation metrics}
In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. Both of them permit to quantify the extent of the Pareto front that is an important factor to consider when evaluating the quality of a Pareto front. Indeed, authors in \cite{zitzler_ec00,weise2012evolutionary} observe that an exploration algorithm should preferably provide Pareto fronts whose objective function range is large. In other words, following the terminology of~\cite{weise2012evolutionary}, a \emph{uniform convergence} should be guaranteed.
Pareto fronts with large objective function ranges permit to investigate in a broader way the potentials of the designed device. This is important, in general, since if only a narrow portion of the Pareto front is provided due to non-uniform convergence, the decision maker may neglect some important options~\cite{weise2012evolutionary}. Having a narrow Pareto front becomes essential, in particular, when designing a device that may be required to work in different operational conditions, for example with stringent execution time constraints and less strict power dissipation ones or, on the contrary, with low power requirements and tolerance to high execution time.

>>>>

\section{A Case Study: VLIW Hardware/Software Design Space}
\secL{ee}
Although PS has been formally described as a general exploration algorithm, the goal of this work is giving a method applicable to the embedded system design. For this reason, we evaluate PS in this context. To this aim, we use a parameterized Very Long Instruction
Word (VLIW) architecture~\cite{kathail_tr00} as testbed. The philosophy
behind a VLIW system makes this a natural choice for several reasons:
\begin{itemize}
\item \emph{Multi-Objective trade-offs}: VLIW architectures allow designers
to trade between power, energy and performances objectives, resulting
in extended Pareto sets which are an ideal ground for our testing
purposes focused on parameter-space representation
\item \emph{Compiler awareness}: code scheduling for the execution of
applications is statically obtained by the compiler which
performs code transformations ``being aware'' of the underlying hardware
configuration.  This tight hardware/software dependence makes the
VLIW scenario perfectly suitable for investigating hardly predictable
parameter interactions.
\end{itemize}

@@

\subsection{Evaluation metrics}
In order to evaluate the Paretos from a quantitative point-of-view we
will consider two metrics, namely, the \emph{variation range} and the
\emph{average normalised absolute dispersion error}. Both of them permit to quantify the extent of the Pareto front that is an important factor to consider when evaluating the quality of a Pareto front. Indeed, authors in \cite{zitzler_ec00,weise2012evolutionary} observe that an exploration algorithm should preferably provide Pareto fronts whose objective function range is large. In other words, following the terminology of~\cite{weise2012evolutionary}, a \emph{uniform convergence} should be guaranteed.
Pareto fronts with large objective function ranges permit to investigate in a broader way the potentials of the designed device. This is important, in general, since if only a narrow portion of the Pareto front is provided due to non-uniform convergence, the decision maker may neglect some important options~\cite{weise2012evolutionary}. Having a wide Pareto front becomes essential, in particular, when designing a device that may be required to satisfy different constraints. For example, the decision maker may want to design two different multi-purpose processors. The first design may adapt to situations in which the processor is used for real time elaboration (thus requiring low execution time) and is attached to the powerline (thus not having stringent power dissipation constraints). The second design, on the contrary, may fit situations in which the energy is provided by a battery (stringent power constraints) but only background computations are performed (long execution time is tolerable). These two processors may be successfully designed maintaining the same architecture, using different values for the design parameters. A wide Pareto front helps in achieving this goal.
